Multi-dimensional Urban Network Percolation

### 2019-10-24 Docs/Papers/Percolation/paper/paper.tex
final_version_paper
prod_version_paper

% header
\title{Multi-dimensional Urban Network Percolation}
% Identifying European Sustainable Mega-city Regions with Multi-dimensional Network Percolation
% Endogenous urban regions as an optimal compromise between economic integration and greenhouse gases emissions
% A multi-dimensional percolation approach to characterize sustainable mega-city regions

@@

Existing heuristics generally focus on a single dimension or property of the urban system. However, such systems are known to be highly multidimensional. For example, the morphological dimension of networks is complementary to the functional properties of the urban environment \citep{burger2012form}. The link between urban form and function remains in particular an open question \citep{batty1994fractal}. More generally, the inclusion of multiple dimension in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \citep{perez2016agent}. This paper addresses such a gap in the case of urban network percolation, by introducing a multi-dimensional percolation heuristic. The method allows combining different dimensions of the urban system, the same way that \cite{cottineau2018defining} link population density with commuting flows to produce multiple definitions of urban areas.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{figures/principle.pdf}}
  \centering
  \caption{Schematic representation of the multi-dimensional network percolation heuristic. Two layers are considered here, with similar nodes but different links in each. To each layer is associated a node variable. In each layer, links are created according to a distance threshold (percolation radius) and a variable value threshold. The final clusters (bottom in color) are obtained by considering the links present in each layer.\label{fig:method}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

We do not call our method ``multi-layer percolation'', since nodes are common between layers. We show in Fig.~\ref{fig:method} for a schematic representation of the method. It can be implemented with a propagation heuristic or directly working on adjacency matrices. The rationale behind combining a thresholding for each layer variable with a distance thresholding relies on the idea that a first component for two points to interact is a large enough proximity. The second component is a strong enough intensity of the urban dimension captured by each layer, simultaneously for all layers considered. This recalls Tobler's first law of geography \citep{tobler2004first} in a multi-dimensional way.

@@

The parameters implied in this heuristic are the percolation radius $r_0$ and the percolation thresholds $\vec{\theta} = \theta_j$ for each layer $j$. Varying these parameters allows considering different levels of percolation.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=0.49\linewidth]{figures/abssize_nodes.png}}
  {\includegraphics[width=0.49\linewidth]{figures/relsize_nodes.png}}
  \centering
  \caption{Percolation transition. On the left, we plot the size of the largest cluster in each configuration in terms of nodes, as a function of the percolation radius $r_0$. Color gives the other percolation parameters. On the right, the plot is similar but with the size relative to the size of the largest cluster obtained with the maximal radius in each configuration.\label{fig:percolation}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{figures/fractaldimension.png}}
  \centering
  \caption{Fractal dimension. We plot for each parametrization given by the curve color the evolution of the fractal dimension $\alpha$ as a function of $r_0$. Standard errors are not plotted for readability.\label{fig:fractaldim}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

In its application to road networks by \cite{arcaute2016cities}, the structure of the national urban system for UK is captured by studying the percolation transition, i.e. the variation of the size of the largest cluster as a function of the percolation radius. As this signature is tightly linked to historical, cultural and geographical conditions, the application to different urban systems should yield different results. We study here this property, for different threshold parameter values. We make the radius vary betweem 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all road network indicators, and three network layer thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$.

@@

% ! largest cluster size in terms of edges makes no sense with the implementation we took - nodes only
%      nwcol         popthq         nwthq          radius         nwparam             fractdim     
% ecount  :186   Min.   :0.85   Min.   :0.00   Min.   :  8000   Length:1023        Min.   :0.9239  
% euclPerf:279   1st Qu.:0.85   1st Qu.:0.00   1st Qu.: 31000   Class :character   1st Qu.:1.2639  
% mu      :279   Median :0.85   Median :0.80   Median : 54000   Mode  :character   Median :1.3306  
% vcount  :279   Mean   :0.85   Mean   :0.55   Mean   : 54000                      Mean   :1.3225  
 %               3rd Qu.:0.85   3rd Qu.:0.95   3rd Qu.: 77000                      3rd Qu.:1.3909  
 %               Max.   :0.85   Max.   :0.95   Max.   :100000                      Max.   :1.6058                                                                                    NA's   :337     
%   fractdimsd     fractdimadjrsquared   fractrelsd    
% Min.   :0.0544   Min.   :0.5111      Min.   :0.0395  
% 1st Qu.:0.0966   1st Qu.:0.7016      1st Qu.:0.0700  
% Median :0.1328   Median :0.7467      Median :0.1043  
% Mean   :0.1317   Mean   :0.7456      Mean   :0.1009  
% 3rd Qu.:0.1578   3rd Qu.:0.7981      3rd Qu.:0.1232  
% Max.   :0.2381   Max.   :0.8642      Max.   :0.1958  
% NA's   :337      NA's   :337         NA's   :337

@@

%  fractdimsignif
%# A tibble: 11 x 5
%# Groups:   nwcol, popthq, nwthq [?]
%   nwcol    popthq nwthq nwparam       signif
%   <fct>     <dbl> <dbl> <chr>          <dbl>
% 1 ecount     0.85  0    ecount0      -0.105 
% 2 ecount     0.85  0.8  ecount0.8    -0.0452
% 3 euclPerf   0.85  0    euclPerf0    -0.108 
% 4 euclPerf   0.85  0.8  euclPerf0.8   0.0360
% 5 euclPerf   0.85  0.95 euclPerf0.95 -0.108 
% 6 mu         0.85  0    mu0          -0.108 
% 7 mu         0.85  0.8  mu0.8        -0.108 
% 8 mu         0.85  0.95 mu0.95        0.375 
% 9 vcount     0.85  0    vcount0      -0.105 
%10 vcount     0.85  0.8  vcount0.8    -0.0416
%11 vcount     0.85  0.95 vcount0.95   -0.0868

%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!] 
%\resizebox{10cm}{7cm}
{\includegraphics[width=0.49\textwidth]{figures/totalPop4183694_00056402_ecount850_radius8000.png}}
  {\includegraphics[width=0.49\linewidth]{figures/totalPop2219780_36719597_vcount378_radius8000.png}}\\
  {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_vcount595_radius50000.png}}
   {\includegraphics[width=0.49\linewidth]{figures/totalPop1474347_36891685_euclPerf0_00507404608502498_radius54000.png}}
  \centering

  \caption{Examples of obtained clusters for different parameter values. In the top-right case for example ($\theta_P = 0.9$, $\theta_N = 0.8$, variable \texttt{vcount},$r_0 = 8km$), we obtain the urban regions of West midlands and London in the UK, Randstad merged with Rhein-Rhur and Rhein-Main in Germany, Paris in France, also with capital cities such as Copenhaguen, Stockholm and Helsinki. There is no cluster in South Europe in that case, due to the high population density threshold.\label{fig:exclusters}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=0.7\linewidth]{figures/full_effective_pareto.png}}
  \centering
  \caption{Point clouds of region-level indicators, namely population and emissions, for different parametrizations, given by the color. Each point represent an endogenous urban region.\label{fig:paretos}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_pc1-emissions_targeted.png}}
  \centering
  \caption{Aggregated values of normalized potential emissions $\sum_c \tilde{g}_c$, as a function of the first morphological principal component (PC1), for varying values of parameters $d_G$ (rows) and $\gamma_G$ (columns). Other intermediate values for these parameters yield similar behaviors. As PC1 is mainly linked to monocentricity, there seems to exist an optimal intermediate level of monocentricity for emissions alone. Color level give the share of population within the considered clusters in comparison to all European population.\label{fig:emissions-pc1}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{figures/aggreg_morpho_relemissions-relefficiency_colpc1_logscale_targeted.png}}
  \centering  
  \caption{Relative potential emissions $\sum_c g_c$ against relative potential economic unefficiency $\sum_c e_c$ (both indicators should be minimized), for varying values of $\gamma_G$ (columns) and $d_G$ (rows). Color level gives the value of PC1, whereas point size gives the share of total population contained within considered areas.\label{fig:paretos-relative}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

%# Cumulative Proportion  0.7296 0.9650 0.99761 1.00000
%                PC1       PC2         PC3         PC4
%moran   -0.3088585 0.9493848 -0.04444327  0.03605266
%avgdist  0.5417362 0.1415668 -0.82239570 -0.10072759
%entropy  0.5108424 0.2140447  0.45847647 -0.69499942
%slope    0.5917502 0.1811415  0.33390034  0.71100630

@@

\bibliographystyle{jimis-en}
\bibliography{biblio}

%\appendix\footnotesize

%\section{Appendix 1: supplementary sensitivity analyses}

%\sframe{Results: effective emissions}{

%\textit{Effective emissions exhibit a supralinear scaling of population}
%\includegraphics[height=0.83\textheight]{figures/aggreg_effective.png}

%\sframe{Results: all clusters Pareto fronts}{

%\textit{Variation of Pareto front patterns when potential parameter $\gamma,d_0$ vary.}

%\includegraphics[width=0.9\textwidth]{figures/full_paretos.png}

%\sframe{Results: an optimal morphology}{

%\textit{More monocentric areas are more optimal in terms of relative emissions and efficiency ?}

%\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relefficiency.png}
%\includegraphics[width=0.49\textwidth]{figures/aggreg_morpho_pc1-relemissions.png}

%  curves for emissions rescale to the same -> link with Caruso profiles ? ; investigate that, and why not for efficiency.

%\section{Acknowledgment}
%

%\section{Biography}
%Here, feel free to add short biographies of the authors.

>>>>

% header
\title{Multi-dimensional Urban Network Percolation}

@@

Existing heuristics generally focus on a single dimension or property of the urban system. However, such systems are known to be highly multidimensional. For example, the morphological dimension of networks is complementary to the functional properties of the urban environment \citep{burger2012form}. The link between urban form and function remains in particular an open question \citep{batty1994fractal}. More generally, the inclusion of multiple dimension in urban analysis is still a research direction to be investigated, as in the case of agent-based models for example \citep{perez2016agent}. This paper addresses such a gap in the case of urban network percolation, by introducing a multi-dimensional percolation heuristic. The method allows combining different dimensions of the urban system, the same way that \cite{cottineau2018defining} link population density with commuting flows to produce multiple definitions of urban areas. We can indeed expect significantly different qualitative behaviors when switching from a single-dimension percolation to a multi-dimensional one, since the spatial structure of different urban dimensions are correlated but also largely complementary. If this is the case, the method then captures more aspects of the urban system and some of their interrelations.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig1.png}}
  \centering
  \caption{Schematic representation of the multi-dimensional network percolation heuristic. Two layers are considered here, with similar nodes but different links in each. To each layer is associated a node variable. In each layer, links are created according to a distance threshold (percolation radius) and a variable value threshold. The final clusters (bottom in color) are obtained by considering the links present in each layer.\label{fig:method}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

We do not call our method ``multi-layer percolation'', since nodes are common between layers. We show in Fig.~\ref{fig:method} for a schematic representation of the method. It can be implemented with a propagation heuristic or directly working on adjacency matrices. The rationale behind combining a thresholding for each layer variable with a distance thresholding relies on the idea that a first component for two points to interact is a geographical proximity. The second component is a strong enough intensity of the urban dimension captured by each layer, simultaneously for all layers considered. This recalls Tobler's first law of geography \citep{tobler2004first} in a multi-dimensional way.

@@

The parameters implied in this heuristic are the percolation radius $r_0$ and the variable thresholds $\vec{\theta} = \theta_j$ for each layer $j$. Varying these parameters allows considering different levels of percolation.

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig2.png}}
  \centering
  \caption{Percolation transition. On the left, we plot the size of the largest cluster in each configuration in terms of nodes, as a function of the percolation radius $r_0$. Color gives the other percolation parameters. On the right, the plot is similar but with the size relative to the size of the largest cluster obtained with the maximal radius in each configuration.\label{fig:percolation}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig3.png}}
  \centering
  \caption{Fractal dimension. We plot for each parametrization given by the curve color the evolution of the fractal dimension $\alpha$ as a function of $r_0$. Standard errors are not plotted for readability.\label{fig:fractaldim}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

In its application to road networks by \cite{arcaute2016cities}, the structure of the national urban system for UK is captured by studying the percolation transition, i.e. the variation of the size of the largest cluster as a function of the percolation radius. As this signature is tightly linked to historical, cultural and geographical conditions, the application to different urban systems should yield different results. We study here this property, for different threshold parameter values. We make the radius vary between 8km and 100km with a one km step, have a fixed population threshold $\theta_P = 0.85$, test all road network indicators, and three network layer thresholds $\theta_N \in \{ 0 ; 0.8 ; 0.95 \}$. These values yield a good precision for the radius which is the most important variable to study transition and estimate fractal dimensions, while this population threshold is enough to provide a broad coverage for large radiuses (as shown by other explorations described below). Changing network variables and their threshold allows us to investigate how the transition behavior do change regarding the dimension considered and its intensity.

@@

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!] 
%\resizebox{10cm}{7cm}
{\includegraphics[width=\linewidth]{Fig4.png}}
  \centering
  \caption{Examples of obtained clusters for different parameter values. In the top-right case for example ($\theta_P = 0.9$, $\theta_N = 0.8$, variable \texttt{vcount},$r_0 = 8km$), we obtain the urban regions of West midlands and London in the UK, Randstad merged with Rhein-Rhur and Rhein-Main in Germany, Paris in France, also with capital cities such as Copenhaguen, Stockholm and Helsinki. There is no cluster in South Europe in that case, due to the high population density threshold.\label{fig:exclusters}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=0.7\linewidth]{Fig5.png}}
  \centering
  \caption{Point clouds of region-level indicators, namely population and emissions, for different parametrizations, given by the color. Each point represent an endogenous urban region.\label{fig:paretos}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig6.png}}
  \centering
  \caption{Aggregated values of normalized potential emissions $\sum_c \tilde{g}_c$, as a function of the first morphological principal component (PC1), for varying values of parameters $d_G$ (rows) and $\gamma_G$ (columns). Other intermediate values for these parameters yield similar behaviors. As PC1 is mainly linked to monocentricity, there seems to exist an optimal intermediate level of monocentricity for emissions alone. Color level give the share of population within the considered clusters in comparison to all European population.\label{fig:emissions-pc1}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht] 
%\resizebox{10cm}{7cm}
  {\includegraphics[width=\linewidth]{Fig7.png}}
  \centering  
  \caption{Relative potential emissions $\sum_c g_c$ against relative potential economic unefficiency $\sum_c e_c$ (both indicators should be minimized), for varying values of $\gamma_G$ (columns) and $d_G$ (rows). Color level gives the value of PC1, whereas point size gives the share of total population contained within considered areas.\label{fig:paretos-relative}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%

@@

@@

\section*{Acknowledgment}

This work is part of DynamiCity, a FUI project funded by BPI France, Auvergne-Rh{\^o}ne-Alpes region, Ile-de-France region and Lyon metropolis.

\begin{thebibliography}{}

\bibitem[\protect\citeauthoryear{Aleta and Moreno}{Aleta and
  Moreno}{2018}]{aleta2018multilayer}
Aleta A., Moreno Y. (2018).
\newblock Multilayer networks in a nutshell.
\newblock {\em Annual Review of Condensed Matter Physics\/}.

\bibitem[\protect\citeauthoryear{Arbabi, Mayfield, and McCann}{Arbabi
  \textit{et~al.}}{2019}]{arbabi2019development}
Arbabi H., Mayfield M., McCann P. (2019).
\newblock On the development logic of city-regions: inter-versus intra-city
  mobility in england and wales.
\newblock {\em Spatial Economic Analysis\/}, 1--20.

\bibitem[\protect\citeauthoryear{Arcaute, Molinero, Hatna, Murcio, Vargas-Ruiz,
  Masucci, and Batty}{Arcaute \textit{et~al.}}{2016}]{arcaute2016cities}
Arcaute E., Molinero C., Hatna E., Murcio R., Vargas-Ruiz C., Masucci A.~P.,
  Batty M. (2016).
\newblock Cities and regions in britain through hierarchical percolation.
\newblock {\em Royal Society open science\/}~{\em 3\/}(4), 150691.

\bibitem[\protect\citeauthoryear{Banos and Genre-Grandpierre}{Banos and
  Genre-Grandpierre}{2012}]{banos2012towards}
Banos A., Genre-Grandpierre C. (2012).
\newblock Towards new metrics for urban road networks: Some preliminary
  evidence from agent-based simulations.
\newblock In {\em Agent-based models of geographical systems}, pp.\  627--641.
  Springer.

\bibitem[\protect\citeauthoryear{Barth{\'e}lemy}{Barth{\'e}lemy}{2011}]{barthelemy2011spatial}
Barth{\'e}lemy M. (2011).
\newblock Spatial networks.
\newblock {\em Physics Reports\/}~{\em 499\/}(1-3), 1--101.

\bibitem[\protect\citeauthoryear{Batty and Longley}{Batty and
  Longley}{1994}]{batty1994fractal}
Batty M., Longley P.~A. (1994).
\newblock {\em Fractal cities: a geometry of form and function}.
\newblock Academic press.

\bibitem[\protect\citeauthoryear{Bettencourt, Lobo, Helbing, K{\"u}hnert, and
  West}{Bettencourt \textit{et~al.}}{2007}]{bettencourt2007growth}
Bettencourt L.~M., Lobo J., Helbing D., K{\"u}hnert C., West G.~B. (2007).
\newblock Growth, innovation, scaling, and the pace of life in cities.
\newblock {\em Proceedings of the national academy of sciences\/}~{\em
  104\/}(17), 7301--7306.

\bibitem[\protect\citeauthoryear{Boccaletti, Bianconi, Criado, Del~Genio,
  G{\'o}mez-Gardenes, Romance, Sendina-Nadal, Wang, and Zanin}{Boccaletti
  \textit{et~al.}}{2014}]{boccaletti2014structure}
Boccaletti S., Bianconi G., Criado R., Del~Genio C.~I., G{\'o}mez-Gardenes J.,
  Romance M., Sendina-Nadal I., Wang Z., Zanin M. (2014).
\newblock The structure and dynamics of multilayer networks.
\newblock {\em Physics Reports\/}~{\em 544\/}(1), 1--122.

\bibitem[\protect\citeauthoryear{Burger and Meijers}{Burger and
  Meijers}{2012}]{burger2012form}
Burger M., Meijers E. (2012).
\newblock Form follows function? linking morphological and functional
  polycentricity.
\newblock {\em Urban studies\/}~{\em 49\/}(5), 1127--1149.

\bibitem[\protect\citeauthoryear{Callaway, Newman, Strogatz, and
  Watts}{Callaway \textit{et~al.}}{2000}]{callaway2000network}
Callaway D.~S., Newman M.~E., Strogatz S.~H., Watts D.~J. (2000).
\newblock Network robustness and fragility: Percolation on random graphs.
\newblock {\em Physical review letters\/}~{\em 85\/}(25), 5468.

\bibitem[\protect\citeauthoryear{Cottineau, Finance, Hatna, Arcaute, and
  Batty}{Cottineau \textit{et~al.}}{2018}]{cottineau2018defining}
Cottineau C., Finance O., Hatna E., Arcaute E., Batty M. (2018).
\newblock Defining urban clusters to detect agglomeration economies.
\newblock {\em Environment and Planning B: Urban Analytics and City Science\/},
  2399808318755146.

\bibitem[\protect\citeauthoryear{Cottineau, Hatna, Arcaute, and
  Batty}{Cottineau \textit{et~al.}}{2017}]{cottineau2017diverse}
Cottineau C., Hatna E., Arcaute E., Batty M. (2017).
\newblock Diverse cities or the systematic paradox of urban scaling laws.
\newblock {\em Computers, environment and urban systems\/}~{\em 63}, 80--94.

\bibitem[\protect\citeauthoryear{Crucitti, Latora, and Porta}{Crucitti
  \textit{et~al.}}{2006}]{crucitti2006centrality}
Crucitti P., Latora V., Porta S. (2006).
\newblock Centrality measures in spatial networks of urban streets.
\newblock {\em Physical Review E\/}~{\em 73\/}(3), 036125.

\bibitem[\protect\citeauthoryear{Ducruet and Beauguitte}{Ducruet and
  Beauguitte}{2014}]{ducruet2014spatial}
Ducruet C., Beauguitte L. (2014).
\newblock Spatial science and network science: Review and outcomes of a complex
  relationship.
\newblock {\em Networks and Spatial Economics\/}~{\em 14\/}(3-4), 297--316.

\bibitem[\protect\citeauthoryear{Feng, Wu, Wu, Su, Weng, and Bian}{Feng
  \textit{et~al.}}{2018}]{feng2018spatiotemporal}
Feng Y., Wu S., Wu P., Su S., Weng M., Bian M. (2018).
\newblock Spatiotemporal characterization of megaregional poly-centrality:
  Evidence for new urban hypotheses and implications for polycentric policies.
\newblock {\em Land use policy\/}~{\em 77}, 712--731.

\bibitem[\protect\citeauthoryear{Florida, Gulden, and Mellander}{Florida
  \textit{et~al.}}{2008}]{florida2008rise}
Florida R., Gulden T., Mellander C. (2008).
\newblock The rise of the mega-region.
\newblock {\em Cambridge Journal of Regions, Economy and Society\/}~{\em
  1\/}(3), 459--476.

\bibitem[\protect\citeauthoryear{Hackett, Cellai, G{\'o}mez, Arenas, and
  Gleeson}{Hackett \textit{et~al.}}{2016}]{hackett2016bond}
Hackett A., Cellai D., G{\'o}mez S., Arenas A., Gleeson J.~P. (2016).
\newblock Bond percolation on multiplex networks.
\newblock {\em Physical Review X\/}~{\em 6\/}(2), 021002.

\bibitem[\protect\citeauthoryear{Hall and Pain}{Hall and
  Pain}{2006}]{hall2006polycentric}
Hall P.~G., Pain K. (2006).
\newblock {\em The polycentric metropolis: learning from mega-city regions in
  Europe}.
\newblock Routledge.

\bibitem[\protect\citeauthoryear{Hillier, Leaman, Stansall, and
  Bedford}{Hillier \textit{et~al.}}{1976}]{hillier1976space}
Hillier B., Leaman A., Stansall P., Bedford M. (1976).
\newblock Space syntax.
\newblock {\em Environment and Planning B: Planning and design\/}~{\em 3\/}(2),
  147--185.

\bibitem[\protect\citeauthoryear{Huynh, Makarov, Legara, Monterola, and
  Chew}{Huynh \textit{et~al.}}{2018}]{huynh2018characterisation}
Huynh H.~N., Makarov E., Legara E.~F., Monterola C., Chew L.~Y. (2018).
\newblock Characterisation and comparison of spatial patterns in urban systems:
  A case study of us cities.
\newblock {\em Journal of computational science\/}~{\em 24}, 34--43.

\bibitem[\protect\citeauthoryear{Janssens-Maenhout, Crippa, Guizzardi, Muntean,
  Schaaf, Dentener, Bergamaschi, Pagliari, Olivier, Peters,
  \textit{et~al.}}{Janssens-Maenhout \textit{et~al.}}{2017}]{janssens2017edgar}
Janssens-Maenhout G., Crippa M., Guizzardi D., Muntean M., Schaaf E., Dentener
  F., Bergamaschi P., Pagliari V., Olivier J., Peters J., \textit{et~al.}
  (2017).
\newblock Edgar v4. 3.2 global atlas of the three major greenhouse gas
  emissions for the period 1970--2012.
\newblock {\em Earth Syst. Sci. Data Discuss\/}.

\bibitem[\protect\citeauthoryear{Komiyama and Takeuchi}{Komiyama and
  Takeuchi}{2006}]{komiyama2006sustainability}
Komiyama H., Takeuchi K. (2006).
\newblock {\em Sustainability science: building a new discipline}.

\bibitem[\protect\citeauthoryear{Lagesse, Bordin, and Douady}{Lagesse
  \textit{et~al.}}{2015}]{lagesse2015spatial}
Lagesse C., Bordin P., Douady S. (2015).
\newblock A spatial multi-scale object to analyze road networks.
\newblock {\em Network Science\/}~{\em 3\/}(1), 156--181.

\bibitem[\protect\citeauthoryear{Lang and Dhavale}{Lang and
  Dhavale}{2005}]{lang2005beyond}
Lang R.~E., Dhavale D. (2005).
\newblock Beyond megalopolis: Exploring america’s new “megapolitan”
  geography.

\bibitem[\protect\citeauthoryear{Laquian}{Laquian}{2011}]{laquian2011planning}
Laquian A. (2011).
\newblock The planning and governance of asia’s mega-urban regions.
\newblock {\em Population Distribution, Urbanization, Internal Migration and
  Development: An International Perspective, United Nations Department of
  Economic and Social Affairs, Population Division, New York: United
  Nations\/}, 302--322.

\bibitem[\protect\citeauthoryear{Lashof and Ahuja}{Lashof and
  Ahuja}{1990}]{lashof1990relative}
Lashof D.~A., Ahuja D.~R. (1990).
\newblock Relative contributions of greenhouse gas emissions to global warming.
\newblock {\em Nature\/}~{\em 344\/}(6266), 529.

\bibitem[\protect\citeauthoryear{Le~N{\'e}chet}{Le~N{\'e}chet}{2012}]{le2012urban}
Le~N{\'e}chet F. (2012).
\newblock Urban spatial structure, daily mobility and energy consumption: a
  study of 34 european cities.
\newblock {\em Cybergeo: European Journal of Geography\/}.

\bibitem[\protect\citeauthoryear{Le~N{\'e}chet}{Le~N{\'e}chet}{2017}]{lenechet2017peupler}
Le~N{\'e}chet F. (2017).
\newblock De l'{\'e}talement urbain aux r{\'e}gions m{\'e}tropolitaines
  polycentriques : formes de fonctionnement et formes de gouvernance.
\newblock In {\em {Peupler la terre - De la pr{\'e}histoire {\`a} l'{\`e}re des
  m{\'e}tropoles}}. Presses Universitaires Francois Rabelais.

\bibitem[\protect\citeauthoryear{Li, Fu, Wang, Lu, Berezin, Stanley, and
  Havlin}{Li \textit{et~al.}}{2015}]{Li669}
Li D., Fu B., Wang Y., Lu G., Berezin Y., Stanley H.~E., Havlin S. (2015).
\newblock Percolation transition in dynamical traffic network with evolving
  critical bottlenecks.
\newblock {\em Proceedings of the National Academy of Sciences\/}~{\em
  112\/}(3), 669--672.
\newblock URL: \url{https://www.pnas.org/content/112/3/669}, \href
  {http://arxiv.org/abs/https://www.pnas.org/content/112/3/669.full.pdf}
  {arXiv:https://www.pnas.org/content/112/3/669.full.pdf}, \href
  {http://dx.doi.org/10.1073/pnas.1419185112} {doi:10.1073/pnas.1419185112}.

\bibitem[\protect\citeauthoryear{Louf and Barthelemy}{Louf and
  Barthelemy}{2013}]{louf2013modeling}
Louf R., Barthelemy M. (2013).
\newblock Modeling the polycentric transition of cities.
\newblock {\em Physical review letters\/}~{\em 111\/}(19), 198702.

\bibitem[\protect\citeauthoryear{Makse, Andrade, Batty, Havlin, Stanley,
  \textit{et~al.}}{Makse \textit{et~al.}}{1998}]{makse1998modeling}
Makse H.~A., Andrade J.~S., Batty M., Havlin S., Stanley H.~E., \textit{et~al.}
  (1998).
\newblock Modeling urban growth patterns with correlated percolation.
\newblock {\em Physical Review E\/}~{\em 58\/}(6), 7054.

\bibitem[\protect\citeauthoryear{Marull, Galletto, Domene, and
  Trull{\'e}n}{Marull \textit{et~al.}}{2013}]{marull2013emerging}
Marull J., Galletto V., Domene E., Trull{\'e}n J. (2013).
\newblock Emerging megaregions: a new spatial scale to explore urban
  sustainability.
\newblock {\em Land Use Policy\/}~{\em 34}, 353--366.

\bibitem[\protect\citeauthoryear{McNally}{McNally}{2000}]{mcnally2000four}
McNally M.~G. (2000).
\newblock The four step model.

\bibitem[\protect\citeauthoryear{Newman and Watts}{Newman and
  Watts}{1999}]{newman1999scaling}
Newman M.~E., Watts D.~J. (1999).
\newblock Scaling and percolation in the small-world network model.
\newblock {\em Physical review E\/}~{\em 60\/}(6), 7332.

\bibitem[\protect\citeauthoryear{Offner, Beaucire, Delaplace, Fr{\'e}mont,
  Ninot, Bretagnolle, and Pumain}{Offner
  \textit{et~al.}}{2014}]{espacegeo2014effets}
Offner J.-M., Beaucire F., Delaplace M., Fr{\'e}mont A., Ninot O., Bretagnolle
  A., Pumain D. (2014).
\newblock Les effets structurants des infrastructures de transport.
\newblock {\em Espace Geographique\/}~(42), p--51.

\bibitem[\protect\citeauthoryear{Penrose}{Penrose}{1999}]{penrose1999k}
Penrose M.~D. (1999).
\newblock On k-connectivity for a geometric random graph.
\newblock {\em Random Structures \& Algorithms\/}~{\em 15\/}(2), 145--164.

\bibitem[\protect\citeauthoryear{Perez, Banos, and Pettit}{Perez
  \textit{et~al.}}{2016}]{perez2016agent}
Perez P., Banos A., Pettit C. (2016).
\newblock Agent-based modelling for urban planning current limitations and
  future trends.
\newblock In {\em International Workshop on Agent Based Modelling of Urban
  Systems}, pp.\  60--69. Springer.

\bibitem[\protect\citeauthoryear{Piovani, Molinero, and Wilson}{Piovani
  \textit{et~al.}}{2017}]{piovani2017urban}
Piovani D., Molinero C., Wilson A. (2017).
\newblock Urban retail location: insights from percolation theory and spatial
  interaction modeling.
\newblock {\em PloS one\/}~{\em 12\/}(10), e0185787.

\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018a}]{raimbault2018calibration}
Raimbault J. (2018a).
\newblock Calibration of a density-based model of urban morphogenesis.
\newblock {\em PloS one\/}~{\em 13\/}(9), e0203516.

\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018b}]{raimbault2018caracterisation}
Raimbault J. (2018b).
\newblock {\em Caract{\'e}risation et mod{\'e}lisation de la co-{\'e}volution
  des r{\'e}seaux de transport et des territoires}.
\newblock Ph.\ D. thesis, Universit{\'e} Paris 7 Denis Diderot.

\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2018c}]{raimbault2018indirect}
Raimbault J. (2018c).
\newblock Indirect evidence of network effects in a system of cities.
\newblock {\em Environment and Planning B: Urban Analytics and City Science\/},
  2399808318774335.

\bibitem[\protect\citeauthoryear{Raimbault}{Raimbault}{2019}]{raimbault2018urban}
Raimbault J. (2019).
\newblock An urban morphogenesis model capturing interactions between networks
  and territories.
\newblock In {\em The Mathematics of Urban Morphology}, pp.\  383--409.
  Springer.

\bibitem[\protect\citeauthoryear{Reuillon, Leclaire, and
  Rey-Coyrehourcq}{Reuillon \textit{et~al.}}{2013}]{reuillon2013openmole}
Reuillon R., Leclaire M., Rey-Coyrehourcq S. (2013).
\newblock Openmole, a workflow engine specifically tailored for the distributed
  exploration of simulation models.
\newblock {\em Future Generation Computer Systems\/}~{\em 29\/}(8), 1981--1990.

\bibitem[\protect\citeauthoryear{Sanders}{Sanders}{2017}]{sanders2017peupler}
Sanders L. (2017).
\newblock {\em Peupler la terre - De la pr{\'e}histoire {\`a} l'{\`e}re des
  m{\'e}tropoles.}
\newblock Presses Universitaires Francois Rabelais.

\bibitem[\protect\citeauthoryear{Son, Bizhani, Christensen, Grassberger, and
  Paczuski}{Son \textit{et~al.}}{2012}]{son2012percolation}
Son S.-W., Bizhani G., Christensen C., Grassberger P., Paczuski M. (2012).
\newblock Percolation theory on interdependent networks based on epidemic
  spreading.
\newblock {\em EPL (Europhysics Letters)\/}~{\em 97\/}(1), 16006.

\bibitem[\protect\citeauthoryear{Sorensen and Okata}{Sorensen and
  Okata}{2010}]{sorensen2010megacities}
Sorensen A., Okata J. (2010).
\newblock {\em Megacities: Urban form, governance, and sustainability},
  Volume~10.
\newblock Springer Science \& Business Media.

\bibitem[\protect\citeauthoryear{Stauffer and Aharony}{Stauffer and
  Aharony}{2014}]{stauffer2014introduction}
Stauffer D., Aharony A. (2014).
\newblock {\em Introduction to percolation theory}.
\newblock Taylor \& Francis.
